A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-04-19 paper/paper1/ms.tex
Add a bunch of text about XUV photoevaporation of hot Saturns

\noindent yielding an incredibly short $<$5 Myr timescale for \hatpb, assuming a typical $\epsilon\sim0.01$.  At Roche lobe overflow, a 5 Myr inflation timescale would imply an instantaneous $\dot{M}_\mathrm{infl}\sim10^{15}$ g/s.  The order of magnitude of this inflation timescale is so fleetingly short that---according to this scenario---\hatpb~must be in the runaway stage of inflation, rapidly losing mass, and growing in surface area to fuel a positive feedback loop.  Under this interpretation, \hatpb~ would represent an example of a new category of planet system that is doomed to evaporate entirely due to the Ohmic Dissipation mechanism, as predicted by \citet{2011ApJ...738....1B}.

@@

\subsection{XUV Photoevaporation}
XUV photoevaporation heats the upper layers in the atmosphere, driving a hydrodynamic wind \citep{2009ApJ...693...23M}.  The combined effects of photoevaporation, anomalous heating, and tidal gravity are predicted to have especially drastic outcomes for inflated hot Saturns \citep{2023ApJ...945L..36T}, where the low densities lead to large mass loss rates.

Photoevaporation stands out as a potential cause of dayside/nightside differences in mass loss, with the greatest supersonic motions stemming from the sub-solar point on the planetary dayside, and merely sonic motion prevailing in latitudinal zonal bands.  The effect may be boosted if dayside/nightside energy transport proceeds inefficiently \citep{2009ApJ...693...23M}.  Such a scenario is illustrated in Figure \ref{fig:KeplerianShear}.

\subsection{Constraining mass loss rate}
A 5 Myr inflation timescale would imply an instantaneous $\dot{M}_\mathrm{infl}\sim10^{15}$ g/s, about 50 times larger than the 1D Parker wind model mass loss rate of $\dot{M}_\mathrm{1D}\sim2\times10^{13}$. Here we enumerate uncertainties that go into assessing mass loss rates to show that some in-between mass loss rate may be plausible.

>>>>

\noindent yielding an incredibly short $<$5 Myr timescale for \hatpb, assuming a typical $\epsilon\sim0.01$.  The order of magnitude of this inflation timescale is so fleetingly short that---according to this scenario---\hatpb~must be in the runaway stage of inflation, rapidly losing mass, and growing in surface area to fuel a positive feedback loop.  Under this interpretation, \hatpb~ would represent an example of a new category of planet system that is doomed to evaporate entirely due to the Ohmic Dissipation mechanism, as predicted by \citet{2011ApJ...738....1B}.  At Roche lobe overflow, a 5 Myr inflation timescale may imply an instantaneous $\dot{M}_\mathrm{infl}>10^{3}$ M$_\oplus$/Gyr.

@@

\subsection{XUV Photoevaporation}
XUV photoevaporation heats the upper layers in the atmosphere, driving a hydrodynamic wind \citep{2009ApJ...693...23M}.  Photoevaporation stands out as offering a natural cause of the observed leading tail attributable to dayside/nightside differences in mass loss: the greatest supersonic motions arise from the sub-solar point on the planetary dayside.  The effect may be boosted if dayside/nightside energy transport proceeds inefficiently \citep{2009ApJ...693...23M}.  Such a scenario is illustrated in Figure \ref{fig:KeplerianShear}.

The combined effects of photoevaporation, anomalous heating, and tidal gravity are predicted to have especially drastic outcomes for inflated hot Saturns \citep{2023ApJ...945L..36T} such as \hatp, where the low densities lead to large mass loss rates:

\begin{equation}
    \dot{M} = \frac{3}{4}\frac{\eta F_\mathrm{XUV}}{G K_t \rho_\mathrm{XUV}}\label{thorn23Mdot}.
\end{equation}

Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ more massive than it is now, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^-3$ threshold, at which point the mass loss rate increases to its current value.  Having lost a significant portion of its mass, its orbit expands to conserve angular momentum.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

\subsection{Constraining mass loss rate}
The theory-inspired mass loss rates above correspond to 1D Parker wind \ion{He}{1} feature depths and widths similar to what we observe, as evinced in Figure \ref{fig:pwinds}, with $\dot{M}_\mathrm{1D}\sim100$ M$_\oplus$/Gyr.  \authorcomment1{To be continued--- interpret this congruence}.
